Higher Moments Distances from Robust Subspace for Botnet Detection

### 2019-05-02 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on literature review (robust pca)

% Anomaly detection by subspace larning
Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}. Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies). Lee \emph{et al.} \cite{Lee2013} presented OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.

% Anomaly detection by robust subspace larning
Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}. The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the problem of decomposing a given data matrix $\textbf{M}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise. This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}. We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for robust a subspace learning review.

RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery. However, RPCA has also been adoted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.

ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map. However, ROBPCA flags many points as outlying when the original data is skewed. To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which improve ROBPCA by means of an adjusted outlyingness based on robust skewness. In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.

Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on robust feature selection and robust PCA for network traffic anomaly detection. It is an unsupervised learning algorithm that allows matching the specificities of the traffic the detector is operating on, which can vary widely both in terms of the most frequent attacks and the relative weight of licit applications.

>>>>

% Anomaly detection by subspace larning
Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}. However, to overcome the above mentioned limitations, some approaches based on PCA have been proposed for anomaly detection. Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies). Lee \emph{et al.} \cite{Lee2013} presented OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.

% Anomaly detection by robust subspace larning
Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}. The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the decomposition of a given data matrix $\textbf{M}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise. This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}. We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.

RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery. However, RPCA has also been adoted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}. ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map. However, ROBPCA flags many points as outlying when the original data is skewed. Therefore, Hubert \emph{et al.} \cite{hubert2009robustskewed} proposed ROBPCA-AO, which improves ROBPCA for problems with skewed data, by means of an adjusted outlyingness based on robust skewness. Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic. The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score distance and the orthogonal distance, calculate the thresholds and classify new observations accordingly.

In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
